\bf ``A p-value of $<$ 0.05 was considered statistically significant'': a content analysis of published statistical methods sections

### 2021-07-20 manuscript/asa_template/asa_template_almostdone.tex
final run of anzctr boilerplate

A similar approach was applied at the sentence level. In this case, we aimed to identify common sentences within statistical methods sections that were most representative of each topic. Analysis therefore focussed
on identifying the reuse of sentences from the top 100 sections assigned to each topic, based on estimated coherence scores. For each example sentence, text was compared with all sections assigned to the same topic.

@@

\begin{landscape}
\begin{table}[]
\centering
\caption{Example boilerplate text from ANZCTR dataset}
\label{tab:anzctr-example-boilerplate}
\begin{tabular}[t]{p{0.1\textwidth}p{0.9\textwidth}p{0.1\textwidth}}
\hline
Topic & Statistical methods text & Matching studies\\
3 & Comparisons between categorical variables will be made either using chi square or Fisher exact test. Continuous data will be compared using the Student’s t-test or Mann-Whitney U test. Two sided p values of less than 0.05 will be considered statistically significant. & \\
 &  The Mann-Whitney U, Student t, 1-way ANOVA, and Kruskal-Wallis tests will be used to compare continuous variables where relevant. The Fisher exact and Pearson’s Chi-square test will be used to compare proportions as appropriate. & \\ \hline
5 & Pilot study & \\
 &  No formal sample size calculation was performed & \\ \hline
7 & Descriptive statistics & \\ \hline
9 & Linear mixed models will be used to analyse the data. & \\ \hline
10 & Repeated measures of ANOVA & \\ \hline
\end{tabular}
\end{table}
\end{landscape}

@@

Statistical software topics differentiated between Prism
GraphPad (topic 3: n = 9974) and SPSS (topic 5: n = 9648).
Within topic 3, targeted searches for ``GraphPad Prism'' identified 7,141 sentences from 6,844 papers. Of these,
1,045 papers contained matching sentences based on our definition of boilerplate text, with instances varying based on version and citation details (Table \ref{tab:plos-example-boilerplate-sentence}).
Similar examples of boilerplate text were identified in topic 5, with ``SPSS'' returning 1,104 matching papers.

@@

\caption{\label{tab:plos-example-boilerplate-sentence} Examples of boilerplate text from PLOS ONE papers based on targeted keyword searches (sentence level). The number of matching papers was based on a Jaccard score of 0.75 or higher}
\centering
\begin{tabular}[t]{p{0.075\textwidth}p{0.15\textwidth}p{0.1\textwidth}p{0.85\textwidth}}
\hline
Topic & Search term &  Matching papers &  Example statistical methods text\\
\hline
\multicolumn{4}{l}{Hypothesis testing}\\
\hline
1 & ``t-test'' & 531 & Statistical analysis was performed using Student's t-test.\\
4 & ``ANOVA'' & 270 & Statistical differences were analysed by one-way ANOVA followed by multiple comparisons performed with post-hoc Bonferroni test (SPSS version 16).\\
9 & ``p $<$'' & 1,962 & $^{*}$p$<$0.05, $^{**}$p$<$0.01, $^{***}$p$<$0.001.\\
\hline
\multicolumn{4}{l}{Statistical software}\\
\hline
3 & ``GraphPad'' &1,786  & All statistical analyses were performed using GraphPad Prism.\\
5 & ``SPSS'' & 1,104 & All statistical analyses were performed using SPSS 17.0 (SPSS Inc Chicago IL USA).\\
\hline
\multicolumn{4}{l}{Descriptive statistics}\\
\hline
6 & ``Mean $\pm$'' & 1,820 & Data are expressed as mean $\pm$ SEM.\\
\hline
\multicolumn{4}{l}{Experiments}\\
\hline
10 & ``Experiment'' & 1,037 & All experiments were repeated at least three times.\\
\hline
\end{tabular}
\end{table}
\end{landscape}

@@

Definitions of statistical significance featured strongly in topic 1 (n
= 3784) and topic 9 (n = 6195). Topic 1 reflected
applications of two-tailed and unpaired Student's t-tests, however,
instances of boilerplate text emphasised the 5\% level of statistical
significance (Table \ref{tab:plos-example-boilerplate}). In contrast, Topic 9
focused on multiple thresholds for declaring statistical significance by
asterisk: ``\(^{*}p<0.05\), \(^{**}p<0.01\) and \(^{***}p<0.001\)'', a
practice that has been criticised \citep{Wasserstein2019}.

Group-based hypothesis testing was a recurring theme across topics, with
text descriptions varying based on method(s) used. One-way analysis of
variance featured strongly in topic 4 (n = 10212), combined
with common methods for performing post-hoc multiple comparisons. Based
on our boilerplate criteria, 1 in 5 studies were matched to the sentence
``Significant differences were determined using analysis of variance
(ANOVA) followed by Tukey post-hoc tests for multiple comparisons''.
Frequently occurring words in topic 6 (n = 4764) reflected
mentions of descriptive statistics for summarising continuous variables,
namely means with standard deviations or standard errors.

Statements of statistical significant was the most frequenty occuring text across nine out of ten topics (\ref{tab:plos-example-boilerplate}). In all cases, statistical significance was defined at p$<$0.05.

>>>>

A similar approach was applied at the sentence level. In this case, we aimed to identify common sentences or phrases within statistical methods sections that were representative of each topic. 
First, keyword searches using dominant words from each topic cloud were completed to identify common variations on sentences involving a keyword. Second, we analysed the reuse of sentences from the top 100 sections assigned to each topic, 
based on estimated coherence scores. For each example sentence, text was compared with all sections assigned to the same topic.

@@

\caption{\label{tab:anzctr-example-boilerplate-sentence} Examples of boilerplate text from ANZCTR studies based on targeted keyword searches (sentence level). The number of matching papers was based on a Jaccard score of 0.75 or higher}
\centering
\begin{tabular}[t]{p{0.075\textwidth}p{0.15\textwidth}p{0.1\textwidth}p{0.85\textwidth}}
\hline
Topic & Search term &  Matching papers &  Example sentence from statistical methods section\\
\hline
\multicolumn{4}{l}{Statistical methods}\\
\hline
3 & ``t-test'' & 131 & Continuous Normally distributed variables will be compared using Student's t-test and reported as means standard deviation while non-normally distributed data will be compared using wilcoxon rank-sum tests and reported as medians inter-quartile-range\\
9 & ``model'' &  339 & Linear mixed models will be used to analyse the data.\\
10 & ``ANOVA'' & 86 & Post-hoc paired comparisons corrected for multiple comparisons will be performed if ANOVA reveal significant effects.\\
\hline
\multicolumn{4}{l}{Study designs}\\
4 & Patient & 375 & At a confidence level of 95\% and a precision around the estimate of 5\% a minimum of 73 patients must be included\\
5 & ``Pilot study'' & & \\
6 & ``Subject' & & \\
8 & ``'group'

\multicolumn{4}{l}{Qualitative analysis}\\
\hline
1 & & &\\
\hline
\multicolumn{4}{l}{Descriptive statistics}\\
\hline
6 & ``Mean $\pm$'' & 1,820 & Data are expressed as mean $\pm$ SEM.\\
\hline

3 & ``GraphPad'' &1,786  & All statistical analyses were performed using GraphPad Prism.\\
5 & ``SPSS'' & 1,104 & All statistical analyses were performed using SPSS 17.0 (SPSS Inc Chicago IL USA).\\
\hline

\multicolumn{4}{l}{Experiments}\\
\hline
10 & ``Experiment'' & 1,037 & All experiments were repeated at least three times.\\
\hline
\end{tabular}
\end{table}
\end{landscape}

\begin{table}
\caption{\label{tab:anzctr-example-boilerplate} Example boilerplate text from ANZCTR studies with the highest number of matches per topic (sentence level). The number of matching to each sentence was based on a Jaccard score of 0.75 or higher and a difference in word count of $\pm$ 3 words; $^{*}$ Most frequent match after excluding statements of statistical significance}
\centering
\begin{tabular}[t]{p{0.05\textwidth}p{0.8\textwidth}p{0.1\textwidth}}
\hline
Topic & Statistical methods text & Matching papers\\
\hline
1 & Significance will be set at p $<$ 0.05 & 13\\
 &  All analyses will be conducted on an intention-to-treat basis$^{*}$ & 8\\
\hline
2 & Analysis will be by intention-to-treat & 15\\
\hline
3 & A p-value $<$ 0.05 is considered statistically significant & 39\\
 & Continuous normally distributed variables will be compared using Student's t-test and reported as means \& standard deviation while non-normally distributed data will be compared using wilcoxon rank-sum tests and reported as medians inter-quartile-range$^{*}$ & 8\\
\hline
4 & Data will be analysed by two methods the Kaplan Meier method and per-protocol analysis & 27\\
\hline
5 & This is a pilot study only & 5\\
\hline
6 & Due to the exploratory nature of this first in human study no power or sample size calculations have been performed & 7\\
\hline
7 & Descriptive statistics & 7\\
\hline
8 & The level of statistical significance will be set at p $<$ 0.05 & 11\\
& Relationships between observations at different occasions will be modelled with an unstructured covariance matrix$^{*}$ & 9\\
\hline
9 & All analyses will be on an intention-to-treat basis & 24\\
\hline
10 & However allowing for a 50 percent attrition rate we will be required to recruit n=70 participants & 10\\

%\begin{landscape}
%\begin{table}[]
%\centering
%\caption{Example boilerplate text from ANZCTR dataset}
%\label{tab:anzctr-example-boilerplate}
%\begin{tabular}[t]{p{0.1\textwidth}p{0.9\textwidth}p{0.1\textwidth}}
%\hline
%Topic & Statistical methods text & Matching studies\\
%3 & Comparisons between categorical variables will be made either using chi square or Fisher exact test. Continuous data will be compared using the Student’s t-test or Mann-Whitney U test. Two sided p values of less than 0.05 will be considered statistically significant. & \\
% &  The Mann-Whitney U, Student t, 1-way ANOVA, and Kruskal-Wallis tests will be used to compare continuous variables where relevant. The Fisher exact and Pearson’s Chi-square test will be used to compare proportions as appropriate. & \\ \hline
%5 & Pilot study & \\
% &  No formal sample size calculation was performed & \\ \hline
%7 & Descriptive statistics & \\ \hline
%9 & Linear mixed models will be used to analyse the data. & \\ \hline
%10 & Repeated measures of ANOVA & \\ \hline
%\end{tabular}
%\end{table}
%\end{landscape}

@@

Statistical software topics differentiated between
GraphPad Prism (topic 3: n = 9,974) and SPSS (topic 5: n = 9,648).
Within topic 3, targeted searches for ``GraphPad'' identified 7,141 sentences from 6,844 papers. Of these,
1,306 papers met our definition of boilerplate text, with examples varying based on version and citation details (Table \ref{tab:plos-example-boilerplate-sentence}).
Similar examples of boilerplate text were identified in topic 5, with ``SPSS'' returning 1,104 matching papers.

@@

\caption{\label{tab:plos-example-boilerplate-sentence} Examples of boilerplate text from PLOS ONE papers based on targeted keyword searches (sentence level). The number of matching papers was based on a Jaccard score of 0.75 or higher}
\centering
\begin{tabular}[t]{p{0.075\textwidth}p{0.15\textwidth}p{0.1\textwidth}p{0.85\textwidth}}
\hline
Topic & Search term &  Matching papers &  Example sentence from statistical methods section\\
\hline
\multicolumn{4}{l}{Hypothesis testing}\\
\hline
1 & ``t-test'' & 531 & Statistical analysis was performed using Student's t-test.\\
4 & ``ANOVA'' & 270 & Statistical differences were analysed by one-way ANOVA followed by multiple comparisons performed with post-hoc Bonferroni test (SPSS version 16).\\
9 & ``p $<$'' & 2,088 & $^{*}$p$<$0.05, $^{**}$p$<$0.01, $^{***}$p$<$0.001.\\
\hline
\multicolumn{4}{l}{Statistical software}\\
\hline
3 & ``GraphPad'' &1,786  & All statistical analyses were performed using GraphPad Prism.\\
5 & ``SPSS'' & 1,104 & All statistical analyses were performed using SPSS 17.0 (SPSS Inc Chicago IL USA).\\
\hline
\multicolumn{4}{l}{Descriptive statistics}\\
\hline
6 & ``Mean $\pm$'' & 1,820 & Data are expressed as mean $\pm$ SEM.\\
\hline
\multicolumn{4}{l}{Experiments}\\
\hline
10 & ``Experiment'' & 1,037 & All experiments were repeated at least three times.\\
\hline
\end{tabular}
\end{table}
\end{landscape}

@@

Definitions of statistical significance featured strongly in topic 1 (n
= 3,784) and topic 9 (n = 6,195). Topic 1 reflected applications of Student's t-tests, however,
instances of boilerplate text emphasised the 5\% level of statistical significance (Table \ref{tab:plos-example-boilerplate}). Topic 9
focused on multiple thresholds for declaring statistical significance by asterisk: ``\(^{*}p<0.05\), \(^{**}p<0.01\) and \(^{***}p<0.001\)'', a
practice that has been criticised \citep{Wasserstein2019}. Variations of this phrase were identfied in 1,676 papers.

Group-based hypothesis testing was a recurring theme across topics, with
text descriptions varying based on method(s) used. One-way analysis of
variance featured strongly in topic 4 (n = 10,212), combined
with common methods for performing post-hoc multiple comparisons (Table \ref{tab:plos-example-boilerplate-sentence}). 
Frequently occurring words in topic 6 (n = 4764) reflected
mentions of descriptive statistics for summarising continuous variables,
namely means with standard deviations or standard errors. The sentence starting with ``data was (were) expressed as (the) mean $\pm$'' returned exact matches in 523 papers.

Statements of statistical significant were the most common form of boilerplate text across nine out of ten topics (\ref{tab:plos-example-boilerplate}). In all cases, statistical significance was defined at p$<$0.05.
